---
output: github_document
---

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

<!-- badges: start -->
[![Lifecycle: maturing](https://img.shields.io/badge/lifecycle-maturing-blue.svg)](https://www.tidyverse.org/lifecycle/#maturing)
[![CRAN status](https://www.r-pkg.org/badges/version/tglkmeans)](https://CRAN.R-project.org/package=tglkmeans)
[![Travis build status](https://travis-ci.com/tanaylab/tglkmeans.svg?branch=master)](https://travis-ci.org/tanaylab/tglkmeans)
[![Codecov test coverage](https://codecov.io/gh/tanaylab/tglkmeans/branch/master/graph/badge.svg)](https://codecov.io/gh/tanaylab/tglkmeans?branch=master)
<!-- badges: end -->


# tglkmeans - efficient implementation of kmeans++ algorithm

This package provides R binding to a cpp implementation of kmeans++ algorithm (https://en.wikipedia.org/wiki/K-means%2B%2B).

## Installation 

You can install the released version of **tglkmeans** using the following command:

```{r, eval=FALSE}
install.packages('tglkmeans', repos=c(getOption('repos'), 'https://tanaylab.github.io/repo'))
```

Or install the develpoment version using: 

```{r, eval=FALSE}
if (!require("remotes")) install.packages("remotes")
remotes::install_github("tanaylab/tglkmeans")
```

## Basic usage
```{r}
library(tglkmeans)
```


Create 5 clusters normally distributed around 1 to 5, with sd of 0.3:

```{r}
data <- rbind(matrix(rnorm(100, mean = 1, sd = 0.3), ncol = 2),
           	  matrix(rnorm(100, mean = 2, sd = 0.3), ncol = 2),
           	  matrix(rnorm(100, mean = 3, sd = 0.3), ncol = 2),
           	  matrix(rnorm(100, mean = 4, sd = 0.3), ncol = 2),
           	  matrix(rnorm(100, mean = 5, sd = 0.3), ncol = 2))
colnames(data) <- c("x", "y")
head(data)
```

Cluster using kmeans++:

```{r}
km <- TGL_kmeans(data, k=5, id_column = FALSE)
km
```

Plot the results:
```{r, clustering, fig.show='hold', fig.asp = 1}
plot(data, col = km$cluster)
points(km$centers, pch=8, cex=2)
```

## Vignette

Please refer to the package vignettes for usage and workflow, or look at the [usage](https://tanaylab.github.io/tglkmeans/articles/usage.html) section in the site.


```{r, eval=FALSE}
browseVignettes('usage') 
```

