---
output: github_document
---

```{r, echo = FALSE}
knitr::opts_chunk$set(
    collapse = TRUE,
    comment = "#>",
    fig.path = "README-"
)
```

<!-- badges: start -->
[![CRAN status](https://www.r-pkg.org/badges/version/tglkmeans)](https://CRAN.R-project.org/package=tglkmeans)
[![Codecov test coverage](https://codecov.io/gh/tanaylab/tglkmeans/branch/master/graph/badge.svg)](https://app.codecov.io/gh/tanaylab/tglkmeans?branch=master)
[![R-CMD-check](https://github.com/tanaylab/tglkmeans/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/tanaylab/tglkmeans/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->


# tglkmeans - efficient implementation of kmeans++ algorithm

This package provides R binding to a cpp implementation of the [kmeans++ algorithm](<https://en.wikipedia.org/wiki/K-means%2B%2B>).

## Installation 

You can install the released version of **tglkmeans** using the following command:

```{r, eval=FALSE}
install.packages("tglkmeans")
```

Or install the development version using: 

```{r, eval=FALSE}
if (!require("remotes")) install.packages("remotes")
remotes::install_github("tanaylab/tglkmeans")
```

## Basic usage
```{r}
library(tglkmeans)
```


Create 5 clusters normally distributed around 1 to 5, with sd of 0.3:

```{r}
data <- rbind(
    matrix(rnorm(100, mean = 1, sd = 0.3), ncol = 2),
    matrix(rnorm(100, mean = 2, sd = 0.3), ncol = 2),
    matrix(rnorm(100, mean = 3, sd = 0.3), ncol = 2),
    matrix(rnorm(100, mean = 4, sd = 0.3), ncol = 2),
    matrix(rnorm(100, mean = 5, sd = 0.3), ncol = 2)
)
colnames(data) <- c("x", "y")
head(data)
```

Cluster using kmeans++:

```{r}
km <- TGL_kmeans(data, k = 5, id_column = FALSE)
km
```

Plot the results:
```{r, clustering, fig.show='hold', fig.asp = 1}
plot(data, col = km$cluster)
points(km$centers, pch = 8, cex = 2)
```

## Vignette

Please refer to the package vignettes for usage and workflow, or look at the [usage](https://tanaylab.github.io/tglkmeans/articles/usage.html) section in the site.


```{r, eval=FALSE}
browseVignettes("usage")
```


## A note regarding random number generation

From version 0.4.0 onwards, the package uses R random number generation functions instead of the C++11 random number generation functions. Note that this may result in different results from previous versions. To get the same results as previous versions, set the `use_cpp_random` argument to `TRUE` in the `TGL_kmeans` function.
