---
title: "basic usage"
author: "Aviezer Lifshitz"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

Basic usage of the package.


## Basic usage

create 5 clusters normally distribution around 1 to 5
```{r}
library(tglkmeans)
data <- purrr::map_df(1:5, ~ 
					as.data.frame(matrix(rnorm(100, mean=.x, sd = 0.3), ncol = 2)) %>%
					mutate(true_clust = .x)) %>% 
				tbl_df %>% 
				mutate(id = 1:n()) %>%
				select(id, everything(), true_clust)
data
```

This is how our true data looks like:

```{r, fig.show='hold'}
data %>% ggplot(aes(x=V1, y=V2, color=factor(true_clust))) + geom_point() + scale_color_discrete(name='true cluster')
```

Now we can cluster it using kmeans++
```{r}
km <- TGL_kmeans_tidy(data %>% select(id, starts_with('V')) , k=5, metric='euclid', verbose=TRUE)
```
The retuned list contains 3 fields:
```{r}
names(km)
```

`km$centers` contains a tibble with `clust` column and the cluster centers:
```{r}
km$centers
```
clusters are numbered according to `order_func` which is hclust by default. 

`km$cluster` contains tibble with `id` column with the observation id (`1:n` if no id column was supplied), and `clust` column with the observation assigned cluster:
```{r}
km$cluster
```

`km$size` contains tibble with `clust` column and `n` column with the number of points in each cluster:
```{r}
km$size
```

We can now check our clustering preformance (fraction of observations that were classified correctly):
```{r}
d <- tglkmeans:::match_clusters(data, km, 5)
sum(d$true_clust == d$new_clust, na.rm=TRUE) / sum(!is.na(d$new_clust))
```

And plot the results:
```{r, fig.show='hold'}
d %>% ggplot(aes(x=V1, y=V2, color=factor(new_clust), shape=factor(true_clust))) + geom_point() + scale_color_discrete(name='cluster') + scale_shape_discrete(name='true cluster')
```

Let's move to higher dimensions (and higher noise):
```{r}
data <- purrr::map_df(1:30, ~ 
					as.data.frame(matrix(rnorm(100*300, mean=.x, sd = 0.7), ncol = 300)) %>%
					mutate(true_clust = .x)) %>% 
				tbl_df %>% 
				mutate(id = 1:n()) %>%
				select(id, everything(), true_clust)
km <- TGL_kmeans_tidy(data %>% select(id, starts_with('V')) , k=30, metric='euclid', verbose=FALSE)
```

```{r}
d <- tglkmeans:::match_clusters(data, km, 30)
sum(d$true_clust == d$new_clust, na.rm=TRUE) / sum(!is.na(d$new_clust))
```
Let's compare it to R vanilla kmeans:

```{r}
km_standard <- kmeans(data %>% select(starts_with('V')), 30)
km_standard$clust <- tibble(id = 1:nrow(data), clust=km_standard$cluster)

d <- tglkmeans:::match_clusters(data, km_standard, 30)
sum(d$true_clust == d$new_clust, na.rm=TRUE) / sum(!is.na(d$new_clust))

```

